\section{Methods as infix operators}
\textnormal{ Methods which take only one parameter can be called as an infix
  operator. It becomes very fun and easy to enchain such methods. With good
  names for objects and methods, a programming instruction can looks like a
  spoken sentence. We add the possibility to call such methods with +, -, *,
  / operators if they are named ''plus'', ''minus'', ''times'' or ''divided''
  respectively.}

\subsection{Example}

\begin{lstlisting}
  class Matrix2 {
    def meaninglessTransformation2() : Matrix2 = {
      var bias1 : Matrix2; bias1.init(1,0,0,1);
      var bias2 : Matrix2; bias2.init(1, 1, 0,0);
      return this.times(bias1).plus(this.times(bias2));  
    }
  }
\end{lstlisting}
\textmd{Becomes}
\begin{lstlisting}
  class Matrix2 {
    def meaninglessTransformation2() : Matrix2 = {
      var bias1 : Matrix2; bias1.init(1,0,0,1);
      var bias2 : Matrix2; bias2.init(1, 1, 0, 0);
      return this * bias1 + this * bias2;
    }
  }
\end{lstlisting}
\subsection{Implementation}
\textnormal{ The first step to implement this feature was to change the 'dot'
  operation definition in the grammars. For the LL1 grammar, we had to add a
  form to the operator, with the possibility of not having a dot, followed by an
  identifier (method) and one single argument with parenthesis or not. }

\begin{lstlisting}
  'OpDot ::= DOT() ~ 'DotEnd ~ 'OpDot
  | 'Identifier ~ 'ExprTerm ~ 'OpDot
  | epsilon(),
  'DotEnd ::= LENGTH()
  | 'Identifier ~ LPAREN() ~ 'Args ~ RPAREN(),
\end{lstlisting}

\textnormal{ To keep our grammar LL1, we had to accept only simple expressions
  ('ExprTerm) as argument in a method call as infix operator, and not composed
  expressions with operator. It seems legit but should be clarified for these
  non-accepted cases : }

\begin{itemize}
\item{Array Read: \colorbox{red}{obj meth arr[index]} is forbidden, we can use instead:\\
    \colorbox{green}{obj.meth(arr[index])} or \colorbox{green}{obj meth
      (arr[index])}}
\item{Bang operation: \colorbox{red}{obj meth !expr} is forbidden, we can use instead:\\
    \colorbox{green}{obj.meth(!expr))} or \colorbox{green}{obj meth (!expr)}}
\item{Dot operation: \colorbox{red}{obj meth obj.meth} is not equivalent to\\
    \colorbox{green}{obj.meth(obj.meth)} or \colorbox{green}{obj meth
      (obj.meth)} but it's equivalent to \colorbox{red}{obj.meth(obj).meth()} \\
    (not forbidden because of the parameterless calls feature that
    we'll describe after.)}
\end{itemize}
\textnormal{A similar and simpler change was also added in the non-LL1 grammar.
  The two constructors had to be adjusted in consequence.}
\textnormal{The operator overloading is then done at type checking and code
  generation stages by allowing operators expression nodes to have Class typed
  operands if the left one has the method named after the operator. The changes
  in the Evaluator follow the same logic. }
